Theories of deliberated choice

### 2024-05-06 simple.tex
False conj about strict claims about types instead of weak ones

Given a set of attitudes $S \wsub P^{\allhists}$, the claim associated to it designates the claim $S^I = \set{{\allleadstog} \in \allleadstos \suchthat \forall i \in I: {\ileadstog} \in S}$.
Say that a claim $C$ is a claim about types iff $\exists S \wsub P^{\allhists} \suchthat C = S^I$.
\begin{definition}
  Say that an attitude $g \in P^{\allhists}$ is observably included in a set of attitudes $S \wsub P^{\allhists}$ iff $\forall H \in \finite(\incr(\allhists)): \exists f \in S \suchthat \restr{g}{H} = \restr{f}{H}$.

@@

  The claim associated to $R({\allleadsto}(I))$ is not trivial: $\forall f \suchthat [\exists \hist, \histp ≠ \emptyset \suchthat \hist_1 = \histp_1 \land f(\hist) ≠ f(\histp)]: f \notin R({\allleadsto}(I))$.
\end{example}
\begin{example}
  Consider $\allargs = \set{\ar_1, \ar_2}$ and ${\allleadsto}(I) = \set{f \suchthat f(\ar_1) ≠ \phi \lor f(\ar_2) ≠ \phi}$, thus, $f \notin {\allleadsto}(I)$ iff $f(\ar_1) = f(\ar_2) = \phi$.

@@

  A function $l: \allhists \to \powerset{P^{\allhists}}$ is decreasing iff $\forall \hist \wsub \histp: l(\histp) \wsub l(\hist)$.
  Let $S \wsub P^{\allhists}$ be a set of attitudes.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-point-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_\hist$.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-sequence-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_{\intvl{\hist}}$.
\end{remark}
\begin{conjecture}
  Given a non empty set of attitudes $S \wsub P^{\allhists}$, the claim associated to $R(S)$ is not trivial iff $\exists \hist \in \allhists, f \in P^{\intvl{\hist}} \suchthat \forall g \in S: \restr{g}{\intvl{hist}} ≠ f$.
\end{conjecture}
\begin{proof}
  This follows from the equivalence of the antecedent with $\exists f \in P^{\allhists}, \hist \in \allhists \suchthat \cap_{\histp \wsub \hist} l^S_f(\histp) = \emptyset$.
\end{proof}
Thus, a situation $S$ does not admit falsifiable theories (meaning that $R(S)$ is associated to the trivial claim) iff $\forall \hist, f, \exists g \in S \suchthat \restr{g}{\intvl{\hist}} = f$.

Generally,
define $S^p = \set{f \in P^{\allhists} \suchthat f(\emptyset) = p}$.
Then \cref{th:caractR} indicates that $R({\allleadsto}(I)) = \set{f \suchthat l^S_f(\emptyset) \in \set{S^p \suchthat p \in P} \land \forall \ar \in \allargs: [\forall \hist \suchthat \hist_1 = \ar: l^S_f(\hist) \in l^S_f(\emptyset)] \land [\forall \hist, \histp \suchthat \hist_1 = \histp_1 = \ar: l^S_f(\hist) = l^S_f(\histp)]}$.
TODO correct this.

\begin{remark}
  The opposite of \cref{th:suffFals} is the negation of $\forall {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Thus, for some ${\allleadstog} \in \allleadstos$ and some $i \in I$ with ${\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}$, we have that $\forall H \in \finite(\incr(\allhists)): \exists j \in I \suchthat \forall \hist \in H: {\ileadstog}(\hist) = {\ileadsto[j]}(\hist)$.

  For example,  and ${\ileadstog}$ is $\phi$ when the path starts with $\ar_1$ (or is empty) and is $¬\phi$ when the path starts with $\ar_2$ and

  In that case, if restricting our focus to claims about types of individuals, we need to include the real types, then close under suitable indistinguishability relation (observable inclusion), so as to include for example $i$.

  Define $S$ as the relations in the claim. Define $R(S)$ as the adequate closure: those that we can’t reject.
  Does $S$ include $R(S)$? If so, it is falsifiable. Otherwise, there is in $R(S)$ a relation that we can’t reject.
  Let $Impl$ be all impliciations in $S$, of the form: if $\hist$ goes to $p$, then $\hist'$ goes to $p'$ (in the above example: if $a$ goes to $\phi$ then $aa$ goes to $\phi$). Any relation that breaks one of these implications is rejectable. For example, sending $a$ to $\phi$ and $ab$ to $¬\phi$ is rejectable.We need to include all relations that break no implication.
  TODO check all this. 
\end{remark}

\begin{theorem}
  Consider the attitude above. Is there a claim that holds and is anonymous and falsifiable?
  Check \cref{ex:falsAnonIfBig}.
\end{theorem}

>>>>

Given a set of attitudes $S \wsub P^{\allhists}$, the claim associated to it designates the claim $S^I = \set{{\allleadstog} \in \allleadstos \suchthat \forall i \in I: {\ileadstog} \in S}$.
Say that a claim $C$ is a claim about types iff $\exists S \wsub P^{\allhists} \suchthat C = S^I$.
Informally, such a claim claims that the set of attitudes in $I$ is contained in $S$ (though might be a strict subset of $S$), or equivalently, that no attitude out of $S$ exists among $I$.
\begin{definition}
  Say that an attitude $g \in P^{\allhists}$ is observably included in a set of attitudes $S \wsub P^{\allhists}$ iff $\forall H \in \finite(\incr(\allhists)): \exists f \in S \suchthat \restr{g}{H} = \restr{f}{H}$.

@@

  The claim associated to $R({\allleadsto}(I))$ is not trivial: $\forall f \suchthat [\exists \hist, \histp \ssub \hist \suchthat f(\histp) ≠ f(\hist)]: f \notin R({\allleadsto}(I))$.
\end{example}
\begin{example}
  Consider $\allargs = \set{\ar_1, \ar_2}$ and ${\allleadsto}(I) = \set{f \suchthat f(\ar_1) ≠ \phi \lor f(\ar_2) ≠ \phi}$, thus, $f \notin {\allleadsto}(I)$ iff $f(\ar_1) = f(\ar_2) = \phi$.

@@

  A function $l: \allhists \to \powerset{P^{\allhists}}$ is decreasing iff $\forall \hist \wsub \histp: l(\histp) \wsub l(\hist)$.
  Let $S \wsub P^{\allhists}$ be a set of attitudes.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-point-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_\hist$.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-sequence-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_{\intvl{\hist}}$.
\end{remark}
\begin{theorem}
  Given $\hist \in \allhists$, define $\restr{S}{\intvl{\hist}} = \set{\restr{f}{\intvl{\hist}} \suchthat f \in S}$.
  Given a non empty set of attitudes $S \wsub P^{\allhists}$, the following statements are equivalent.
  \begin{itemize}
    \item The claim associated to $R(S)$ is not trivial 
    \item $R(S) ≠ P^{\allhists}$
    \item $\exists \hist \in \allhists, f \in P^{\intvl{\hist}} \suchthat \forall g \in S: \restr{g}{\intvl{hist}} ≠ f$
    \item $\exists \hist, f \in P^{\intvl{\hist}} \suchthat f \notin \restr{S}{\intvl{\hist}}$
    \item $\exists \hist \suchthat P^{\intvl{\hist}} \not\wsub \restr{S}{\intvl{\hist}}$.
  \end{itemize}
  Also, the following statements are equivalent.
  \begin{itemize}
    \item The claim associated to $R(S)$ is trivial
    \item $R(S) = P^{\allhists}$
    \item $\forall \hist, f, \exists g \in S \suchthat \restr{g}{\intvl{\hist}} = f$
    \item $\forall \hist: P^{\intvl{\hist}} \wsub \restr{S}{\intvl{\hist}}$.
  \end{itemize}
  Furthermore, $\forall \hist: P^{\intvl{\hist}} \wsub \restr{S}{\intvl{\hist}} \implies \card{S} \notin \N$.
\end{theorem}
\begin{proof}
  This follows from $R(S) ≠ P^{\allhists} \iff \exists f \in P^{\allhists}, \hist \in \allhists \suchthat \cap_{\histp \wsub \hist} l^S_f(\histp) = \emptyset$.
\end{proof}

To recap, we now know that with $S = {\allleadsto}(I)$, when $R(S) ≠ P^{\allhists}$, there is a claim that holds and is anonymous and falsifiable.
This includes all situations with finite behaviors ($\card{S} \in \N$), thus includes all situations with finite $I$.
Consider now situations such that $R({\allleadsto}(I)) = P^{\allhists}$.

Define $I = P^\N$ and ${\ileadsto} = i \circ \card$, so that $S = {\allleadsto}(I) = \set{l \circ \card \in P^{\allhists} \suchthat l \in P^\N}$.
We have $R(S) = P^{\allhists}$.
The function $f(\ar) ≠ f(\ar')$, for example, is not in $S$.

\begin{conjecture}
  Consider a mapping ${\allleadstog} \in (P^{\allhists})^I$. Define $E$ as the equivalence classes in $I$ that considers as equivalent all $i$ mapped to the same element of $P^{\allhists}$.
  A wide-shuffling over $I$ is a mapping $\sigma_{\allleadstog} \in I^I$ that reaches any equivalence class in $E$, thus $\forall e \in E: \exists i \suchthat \sigma_{\allleadstog}(i) \in e$.
  The bijections over $I^I$ are a subset of the wide-shufflings (those that also are bijections per $e \in E$, thus $\forall e \in E: \restr{\sigma_{\allleadstog}}{e}$ bijects with its image).

  A claim $C$ is not about types iff $\exists {\allleadstog} \in C$ and $\sigma_{\allleadstog}$, a wide-shuffling, such that $({\allleadstog} \circ \sigma_{\allleadstog}) \notin C$.
\end{conjecture}
\begin{proof}
  % Consider that $\exists {\allleadstog} \in C$ and $\sigma_{\allleadstog}$, a wide-shuffling, such that $({\allleadstog} \circ \sigma_{\allleadstog}) \notin C$.
  Consider a claim $C$ about types $S$.
  Consider any ${\allleadstog} \in C$ and $\sigma_{\allleadstog}$, a wide-shuffling.
  As ${\allleadstog} \in C$, we have ${\allleadstog}(I) \wsub S$.
  Also, ${\allleadstog}(\sigma_{\allleadstog}(I)) = {\allleadstog}(I)$. 
  Thus, ${\allleadstog}(\sigma_{\allleadstog}(I)) \wsub S$.
  Thus, $({\allleadstog} \circ \sigma_{\allleadstog}) \in C$.

  About the converse, given $C$, 
  define $S = \bigcup_{{\allleadstog} \in C}({\allleadstog}(I))$,
  and assuming that 
  $[\forall {\allleadstog} \in C, \sigma_{\allleadstog}: ({\allleadstog} \circ \sigma_{\allleadstog}) \in C]$, let us show that 
  $C = S^I$, equivalently, that $C = \set{{\allleadstog} \in (P^{\allhists})^I \suchthat {\allleadstog}(I) \wsub S}$.
  First, considering any ${\allleadstog} \in C$, we have ${\allleadstog}(I) \wsub S$ by definition of $S$.
  Second, considering any ${\allleadstog} \notin C$, 
\end{proof}
\begin{conjecture}
  Consider a claim that is not about types but is anonymous. What’s that?
  Example: the claim “everyone has the same behavior”.

  Example 2, with $\card{I} = 3$: “exactly two individuals have behavior 1”.

  Is there an example with infinitely many behaviors?
  The claim must be not about types. If non empty, the claim includes some mapping in $(P^{\allhists})^I$. If not about types, some such mapping has not all correspondances shuffling $I$ in any number but keeping the same range?

  Is it that if there is a claim anonymous, holding, and fals., then there is a claim about types that is anonymous, holds, and fals?
\end{conjecture}
\begin{theorem}
  Consider the attitude above. Is there a claim that holds and is anonymous and falsifiable?
  Check \cref{ex:falsAnonIfBig}.
\end{theorem}
